---
title: "Full Pipeline: FD scRNA-seq (GSE301458) - Osteoprogenitor FD vs CTRL"
---

Pipeline stages:
1. Download & organize GEO data
2. Per-sample QC + SoupX ambient RNA removal
3. Doublet removal
4. Merge, normalize, integrate
5. Cluster + annotate cell types
6. Subset osteoprogenitor cells
7. Differential expression: FD vs CTRL (gene list, geomean/group, FC, log2FC, p-val, padj)

IMPORTANT CAVEAT: this scRNA count matrix does NOT contain per-cell GNAS genotype calls.
Those come from a separate targeted-PCR/BaseScope genotyping dataset in the paper. What
this pipeline gives you is the FD-tissue vs CTRL-tissue transcriptomic comparison within
the osteoprogenitor cluster (the "FD-specific osteoblastic cluster" the paper describes),
not a mutant-cell-vs-wildtype-cell comparison within the same sample. If you want the
latter, you need to merge in the genotype-calls file (check GEO supplementary files /
paper's supplementary tables for a per-barcode genotype table) and add a genotype column
to metadata before running FindMarkers.

Run each chunk stage by stage and inspect intermediate output (QC plots, UMAPs, marker
dotplots) before moving to the next stage - don't run this end-to-end blind.

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r libraries}
library(Seurat)      # v5
library(SoupX)
library(scDblFinder)  # doublet detection (Bioconductor)
library(SingleCellExperiment)
library(harmony)      # integration
library(GEOquery)
library(Matrix)
library(dplyr)
library(tibble)
library(tidyr)
library(ggplot2)
library(stringr)
library(DESeq2)
```

# ============================================================
# STAGE 1: Download and organize data
# ============================================================

```{r download}
dir.create("GSE301458", showWarnings = FALSE)
getGEOSuppFiles("GSE301458", makeDirectory = TRUE, baseDir = ".")
untar("GSE301458/GSE301458_RAW.tar", exdir = "GSE301458/raw")
raw_dir <- "GSE301458/raw"
list.files(raw_dir) |> head(30)
```

GEO 10x deposits are usually named like:
`GSMxxxxxxx_<samplename>_barcodes.tsv.gz`, `..._features.tsv.gz`, `..._matrix.mtx.gz`
Run the listing above FIRST and inspect actual filenames before trusting the parsing
below - GEO submitters are inconsistent, so adjust the regex to match what you see.

```{r organize_per_sample}
all_files <- list.files(raw_dir, full.names = TRUE)

# Only pick up the 10x triplet files (barcodes/features/matrix) - exclude the
# separate GNAS genotype csv files, which are handled separately below.
tenx_files <- all_files[str_detect(basename(all_files), "_(barcodes|features|genes|matrix)\\.(tsv|mtx)\\.gz$")]

# Pull out unique sample prefixes (everything before _barcodes/_features/_matrix)
sample_prefixes <- tenx_files |>
  basename() |>
  str_replace("_(barcodes|features|genes|matrix)\\.(tsv|mtx)(\\.gz)?$", "") |>
  unique()

sample_prefixes

# GNAS genotype files (FD samples only) - map sample prefix -> genotype csv path
genotype_files <- all_files[str_detect(basename(all_files), "gnasGTcounts\\.csv\\.gz$")]
names(genotype_files) <- basename(genotype_files) |>
  str_replace("_gnasGTcounts\\.csv\\.gz$", "")
genotype_files

# IMPORTANT: the RNA count matrix and the GNAS genotype file for the SAME biological
# sample were assigned DIFFERENT GSM accession numbers by GEO, e.g.:
#   GSM9083544_FD16_barcodes.tsv.gz          (RNA counts)
#   GSM9083548_FD16_gnasGTcounts.csv.gz      (genotype calls)
# Matching on the full prefix (including GSM number) will fail to pair these up, so we
# re-key genotype_files by the bare sample name only (e.g. "FD16") here, and will do the
# same to sample_metadata$prefix once it's created in the next chunk.

extract_sample_name <- function(x) str_extract(x, "(FD|WT)\\d+")

names(genotype_files) <- extract_sample_name(names(genotype_files))
genotype_files   # keys should now just be "FD16", "FD21", "FD28", "FD716"
```

```{r make_10x_dirs}
# Read10X() needs a directory per sample containing exactly:
# barcodes.tsv.gz, features.tsv.gz (or genes.tsv.gz), matrix.mtx.gz
# So we symlink/copy each sample's 3 files into its own folder with standard names.

dir.create("GSE301458/per_sample", showWarnings = FALSE)

for (pfx in sample_prefixes) {
  sample_dir <- file.path("GSE301458/per_sample", pfx)
  dir.create(sample_dir, showWarnings = FALSE)

  bc  <- tenx_files[str_detect(basename(tenx_files), paste0("^", pfx, "_barcodes"))]
  ft  <- tenx_files[str_detect(basename(tenx_files), paste0("^", pfx, "_(features|genes)"))]
  mx  <- tenx_files[str_detect(basename(tenx_files), paste0("^", pfx, "_matrix"))]

  if (length(bc) == 1 && length(ft) == 1 && length(mx) == 1) {
    file.copy(bc, file.path(sample_dir, "barcodes.tsv.gz"), overwrite = TRUE)
    file.copy(ft, file.path(sample_dir, "features.tsv.gz"), overwrite = TRUE)
    file.copy(mx, file.path(sample_dir, "matrix.mtx.gz"), overwrite = TRUE)
  } else {
    warning("Unexpected file count for sample: ", pfx)
  }
}
```

```{r assign_groups}
# Confirmed directly from GSE301458 raw file listing:
#   GSM9083540_WT150, GSM9083541_WT151, GSM9083542_WT152, GSM9083543_WT153  -> CTRL
#   GSM9083544_FD16,  GSM9083545_FD21,  GSM9083546_FD28,  GSM9083547_FD716  -> FD
#   (matches the sample names used in the lab's own SoupX preprocessing script)

sample_metadata <- tibble(prefix = sample_prefixes) %>%
  mutate(
    sample_name = extract_sample_name(prefix),  # e.g. "FD16", "WT150" - used to key
                                                  # genotype files (see organize_per_sample
                                                  # chunk - GSM accessions differ between
                                                  # the RNA matrix and genotype files for
                                                  # the same sample, so we can't match on
                                                  # the full prefix)
    group = case_when(
      str_detect(prefix, "WT")  ~ "CTRL",
      str_detect(prefix, "FD")  ~ "FD",
      TRUE ~ NA_character_
    )
  )

sample_metadata
stopifnot(!any(is.na(sample_metadata$group)))  # fail loudly if anything unmapped
```

# ============================================================
# STAGE 2: Per-sample load, QC, SoupX ambient RNA removal
# ============================================================

SoupX ideally wants BOTH the raw (unfiltered) and filtered matrices per sample to
estimate ambient contamination. If GEO only provides filtered (cell-called) matrices,
SoupX can still run using the filtered matrix as a proxy but contamination estimation
is less accurate - check what GEO actually gives you.

```{r load_qc_soupx}
seurat_list <- list()

for (i in seq_len(nrow(sample_metadata))) {
  pfx    <- sample_metadata$prefix[i]
  sname  <- sample_metadata$sample_name[i]   # e.g. "FD16" - use this to key genotype_files
  group  <- sample_metadata$group[i]
  if (is.na(group)) next  # skip unmapped samples

  sample_dir <- file.path("GSE301458/per_sample", pfx)
  counts <- Read10X(data.dir = sample_dir)

  so <- CreateSeuratObject(counts = counts, project = pfx, min.cells = 3, min.features = 200)
  so$sample  <- pfx
  so$group   <- group
  so[["percent.mt"]] <- PercentageFeatureSet(so, pattern = "^MT-")

  # ---- GNAS genotype integration (FD samples only) ----
  if (sname %in% names(genotype_files)) {
    gt <- read.csv(genotype_files[[sname]])
    # gt columns: Cell_Barcode, GNAS_Genotype, read_count
    # Raw GNAS_Genotype values seen in this dataset: "601.CtoT", "601.CtoT.pA", "WT", "WT.pA"
    # 601.CtoT = c.601C>T coding change (GNAS R201H) = mutant; .pA suffix is a technical
    # read-type variant, not a distinct genotype, so both variants collapse together.
    gt <- gt %>%
      mutate(GNAS_call = case_when(
        str_detect(GNAS_Genotype, "CtoT") ~ "MUT",
        str_detect(GNAS_Genotype, "WT")   ~ "WT",
        TRUE ~ NA_character_
      ))

    geno_vec  <- gt$GNAS_call[match(colnames(so), gt$Cell_Barcode)]
    reads_vec <- gt$read_count[match(colnames(so), gt$Cell_Barcode)]

    so$gnas_genotype   <- geno_vec       # NA for cells with no genotyping coverage
    so$gnas_read_count <- reads_vec

    n_called <- sum(!is.na(geno_vec))
    message(pfx, ": ", n_called, " / ", ncol(so), " cells have a GNAS genotype call (",
            sum(geno_vec == "MUT", na.rm = TRUE), " MUT, ",
            sum(geno_vec == "WT", na.rm = TRUE), " WT)")
  } else {
    so$gnas_genotype   <- NA_character_  # CTRL samples: no targeted genotyping done
    so$gnas_read_count <- NA_integer_
  }

  seurat_list[[pfx]] <- so
}

# QC snapshot before filtering - INSPECT before choosing thresholds
qc_df <- purrr::map_dfr(seurat_list, function(x) x@meta.data, .id = "sample")
ggplot(qc_df, aes(x = sample, y = nFeature_RNA)) + geom_violin() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggplot(qc_df, aes(x = sample, y = percent.mt)) + geom_violin() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r qc_filter}
# ---- USER-EDITABLE thresholds - set based on the violin plots above ----
min_features <- 200
max_features <- 6000
max_pct_mt   <- 15

seurat_list <- lapply(seurat_list, function(so) {
  subset(so, subset = nFeature_RNA > min_features &
                       nFeature_RNA < max_features &
                       percent.mt   < max_pct_mt)
})
```

```{r doublets}
# Doublet detection per sample (do this BEFORE merging, per-sample, as recommended)
seurat_list <- lapply(seurat_list, function(so) {
  sce <- as.SingleCellExperiment(so)
  sce <- scDblFinder(sce)
  so$scDblFinder.class <- sce$scDblFinder.class
  subset(so, subset = scDblFinder.class == "singlet")
})
```

# ============================================================
# STAGE 3: Merge, normalize, integrate
# ============================================================

```{r merge}
merged <- merge(seurat_list[[1]], y = seurat_list[-1], add.cell.ids = names(seurat_list))
merged <- JoinLayers(merged)  # Seurat v5
```

```{r normalize_cluster}
merged <- NormalizeData(merged)
merged <- FindVariableFeatures(merged, nfeatures = 2000)
merged <- ScaleData(merged)
merged <- RunPCA(merged, npcs = 30)

# Harmony integration across samples (batch = sample, adjust if you also want to
# regress out chemistry/platform, since this study used both MiSeq and NovaSeq)
merged <- RunHarmony(merged, group.by.vars = "sample")

merged <- RunUMAP(merged, reduction = "harmony", dims = 1:30)
merged <- FindNeighbors(merged, reduction = "harmony", dims = 1:30)
merged <- FindClusters(merged, resolution = 0.6)

DimPlot(merged, group.by = "seurat_clusters", label = TRUE)
DimPlot(merged, group.by = "group")
DimPlot(merged, group.by = "sample")
```

# ============================================================
# STAGE 4: Cell type annotation - identify osteoprogenitor cluster
# ============================================================

Canonical osteoprogenitor/osteoblast-lineage markers (per Wentworth et al. and standard
skeletal stromal biology): RUNX2, SP7 (Osterix), ALPL, COL1A1, SPP1, IBSP, BGLAP.
Also check endothelial (PECAM1, VWF), perivascular/pericyte (RGS5, PDGFRB), and general
fibroblast (PDGFRA, DCN) markers to make sure you're not mis-calling clusters, since the
paper specifically highlights overlap of GNAS expression across stromal, endothelial, and
perivascular lineages.

```{r marker_dotplot}
markers <- c(
  "RUNX2", "SP7", "ALPL", "COL1A1", "SPP1", "IBSP", "BGLAP",  # osteoprogenitor/osteoblast
  "PDGFRA", "DCN", "LUM",                                      # fibroblast
  "PECAM1", "VWF",                                             # endothelial
  "RGS5", "PDGFRB", "MYH11"                                    # perivascular/pericyte
)

DotPlot(merged, features = markers, group.by = "seurat_clusters") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r annotate}
# Assigned from the marker dotplot data (see chat/notes for full per-cluster rationale).
# Clusters 6, 11, 14, 15, 16, 17 are excluded: 11/14/16/17 are PTPRC+ (immune/
# hematopoietic contamination - out of scope per the paper's non-hematopoietic focus),
# and 6/15 lack a clear/consistent marker signature (6 = no elevated markers at all;
# 15 = simultaneously high chondrocyte AND neural markers, likely low-quality/doublets).
cluster_to_celltype <- c(
  "0"  = "Fibroblast",
  "1"  = "Endothelial",
  "2"  = "Perivascular",
  "3"  = "Fibroblast",
  "4"  = "Osteoprogenitor",              # <-- primary population of interest
  "5"  = "Fibroblast",
  "6"  = NA_character_,                  # unclear/low-quality - excluded
  "7"  = "Endothelial",
  "8"  = "Fibroblast",
  "9"  = "Osteoprogenitor_primed",       # SPP1/IBSP/BGLAP-positive but RUNX2/ALPL-low;
                                          # kept separate from cluster 4, not merged, since
                                          # it's a weaker/less mature osteoblast-lineage
                                          # signature - decide later whether to pool these
  "10" = "Endothelial",                  # sinusoidal/CXCL12+LEPR+ subtype
  "11" = NA_character_,                  # PTPRC+ immune - excluded
  "12" = "Adipocyte_stromal",            # LepR+/Adipoq+ marrow stromal/MSC
  "13" = "Osteoclast",
  "14" = NA_character_,                  # PTPRC+ macrophage-like - excluded
  "15" = NA_character_,                  # mixed chondro/neural signature - excluded
  "16" = NA_character_,                  # weak PTPRC+, low-quality - excluded
  "17" = NA_character_                   # weak PTPRC+, low-quality - excluded
)

merged$celltype <- unname(cluster_to_celltype[as.character(merged$seurat_clusters)])
table(merged$celltype, useNA = "always")
table(merged$celltype, merged$group)
table(merged$sample[merged$celltype == "Osteoprogenitor"], merged$group[merged$celltype == "Osteoprogenitor"])
table(merged$sample)  # total cells captured per sample, post-QC, all cell types
```

```{r save_checkpoint}
dir.create("Objects", showWarnings = FALSE)
saveRDS(merged, "Objects/FD_integrated_annotated.rds")
```

# ============================================================
# STAGE 5: Osteoprogenitor DE: FD vs CTRL
# ============================================================

```{r subset_osteoprog}
op <- subset(merged, subset = celltype == "Osteoprogenitor")
op <- subset(op, subset = group %in% c("FD", "CTRL"))
Idents(op) <- op$group
table(Idents(op))
```

```{r find_markers}
de <- FindMarkers(
  op,
  ident.1 = "FD", ident.2 = "CTRL",
  test.use = "wilcox",
  logfc.threshold = 0, min.pct = 0, min.cells.group = 0
) %>%
  rownames_to_column("gene")
```

```{r geomean_fc}
pseudocount <- 1
expr <- GetAssayData(op, assay = "RNA", layer = "data")

cells_fd   <- WhichCells(op, idents = "FD")
cells_ctrl <- WhichCells(op, idents = "CTRL")

geomean_fd   <- Matrix::rowMeans(expm1(expr[, cells_fd,   drop = FALSE]))
geomean_ctrl <- Matrix::rowMeans(expm1(expr[, cells_ctrl, drop = FALSE]))
pct_fd       <- Matrix::rowMeans(expr[, cells_fd,   drop = FALSE] > 0)
pct_ctrl     <- Matrix::rowMeans(expr[, cells_ctrl, drop = FALSE] > 0)

geo_df <- tibble(
  gene = names(geomean_fd),
  geomean_FD = geomean_fd, geomean_CTRL = geomean_ctrl,
  pct_FD = pct_fd, pct_CTRL = pct_ctrl
) %>%
  mutate(FC = (geomean_FD + pseudocount) / (geomean_CTRL + pseudocount),
         log2FC = log2(FC))
```

```{r combine_results}
results <- geo_df %>%
  left_join(de %>% select(gene, p_val, p_val_adj), by = "gene") %>%
  arrange(p_val_adj, desc(abs(log2FC)))

results %>% filter(gene == "GNAS")

dir.create("Results", showWarnings = FALSE)
write.csv(results, "Results/Osteoprogenitor_FD_vs_CTRL_DE.csv", row.names = FALSE)
```

```{r volcano}
results %>%
  mutate(neglog10padj = -log10(p_val_adj + 1e-300)) %>%
  ggplot(aes(x = log2FC, y = neglog10padj)) +
  geom_point(alpha = 0.4, size = 0.8) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  geom_vline(xintercept = c(-0.25, 0.25), linetype = "dashed") +
  labs(title = "Osteoprogenitor cells: FD vs CTRL",
       x = "log2FC (FD / CTRL)", y = "-log10(adj p-value)") +
  theme_minimal()
```

# ============================================================
# STAGE 6: The comparison you actually want - GNAS-MUTANT vs WT
# osteoprogenitor cells (within FD samples, using real per-cell genotype)
# ============================================================

This is the mosaicism-resolved comparison: cells within the SAME FD patients that carry
the GNAS mutation vs cells that don't, restricted to the osteoprogenitor population.
This directly tests "does the mutation itself affect the osteoblast progenitor
transcriptome" rather than "is FD tissue different from control tissue" (which could be
confounded by anything else different about FD patients/samples, e.g. age, biopsy site).

```{r subset_by_genotype}
# ---- USER-EDITABLE: minimum reads to trust a genotype call ----
# Single-UMI/read calls are noisier (higher chance of a stray mis-called base). Consider
# requiring >= 2-3 reads before trusting a call - inspect the read_count distribution
# first if you want to tune this rather than accepting the default below.
min_gnas_reads <- 2

merged$gnas_genotype_filtered <- ifelse(
  !is.na(merged$gnas_genotype) & merged$gnas_read_count >= min_gnas_reads,
  merged$gnas_genotype,
  NA_character_
)

# Only FD-sample osteoprogenitor cells with a confident genotype call
op_fd <- subset(merged, subset = celltype == "Osteoprogenitor" & group == "FD")

table(op_fd$gnas_genotype_filtered, useNA = "always")

op_fd_called <- subset(op_fd, subset = !is.na(gnas_genotype_filtered))
Idents(op_fd_called) <- op_fd_called$gnas_genotype_filtered
table(Idents(op_fd_called))  # check both groups have enough cells (rule of thumb: ideally >=10-20 per group for a stable Wilcoxon test)
```

```{r geno_de}
geno_de <- FindMarkers(
  op_fd_called,
  ident.1 = "MUT", ident.2 = "WT",
  test.use = "wilcox",
  logfc.threshold = 0, min.pct = 0, min.cells.group = 0
) %>%
  rownames_to_column("gene")

expr2 <- GetAssayData(op_fd_called, assay = "RNA", layer = "data")
cells_mut <- WhichCells(op_fd_called, idents = "MUT")
cells_wt  <- WhichCells(op_fd_called, idents = "WT")

geomean_mut <- Matrix::rowMeans(expm1(expr2[, cells_mut, drop = FALSE]))
geomean_wt  <- Matrix::rowMeans(expm1(expr2[, cells_wt,  drop = FALSE]))
pct_mut     <- Matrix::rowMeans(expr2[, cells_mut, drop = FALSE] > 0)
pct_wt      <- Matrix::rowMeans(expr2[, cells_wt,  drop = FALSE] > 0)

geno_results <- tibble(
  gene = names(geomean_mut),
  geomean_MUT = geomean_mut, geomean_WT = geomean_wt,
  pct_MUT = pct_mut, pct_WT = pct_wt
) %>%
  mutate(FC = (geomean_MUT + pseudocount) / (geomean_WT + pseudocount),
         log2FC = log2(FC)) %>%
  left_join(geno_de %>% select(gene, p_val, p_val_adj), by = "gene") %>%
  arrange(p_val_adj, desc(abs(log2FC)))

geno_results %>% filter(gene == "GNAS")  # sanity check - not expected to show
                                          # differential GNAS *expression* necessarily,
                                          # since this is a point mutation not a
                                          # expression-level change

write.csv(geno_results, "Results/Osteoprogenitor_GNASmut_vs_WT_DE.csv", row.names = FALSE)
```

```{r geno_volcano}
geno_results %>%
  mutate(neglog10padj = -log10(p_val_adj + 1e-300)) %>%
  ggplot(aes(x = log2FC, y = neglog10padj)) +
  geom_point(alpha = 0.4, size = 0.8) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  geom_vline(xintercept = c(-0.25, 0.25), linetype = "dashed") +
  labs(title = "Osteoprogenitor cells: GNAS-MUT vs WT (within FD samples, per-cell test)",
       x = "log2FC (MUT / WT)", y = "-log10(adj p-value)") +
  theme_minimal()
```

# ============================================================
# STAGE 7: Paired pseudobulk DESeq2 - the more defensible comparison
# ============================================================

The per-cell Wilcoxon test above treats every cell as an independent replicate, but
cells from the same patient are correlated (pseudoreplication) - with mutant calls this
skewed across patients (FD16: 55 MUT / 350 WT, FD28: 7 MUT / 475 WT, FD21 & FD716: 0 MUT
even unfiltered), the per-cell test above is largely a FD16-vs-rest comparison wearing a
genotype label, not a clean genotype effect. Pseudobulk aggregation to sample x genotype,
paired by patient, is the standard fix (Squair et al. 2021, Nat Commun) - it collapses
each patient x genotype group into one pseudo-sample, so genotype is tested at the
correct level of biological replication (patient), with patient identity explicitly
included as a blocking factor.

Only FD16 and FD28 have both MUT and WT osteoprogenitor cells, so this is necessarily a
paired, 2-patient analysis - treat results as hypothesis-generating (n=2 pairs is not
enough for strong statistical claims), not as confirmatory as a properly powered cohort
would give you. Report this limitation explicitly in any write-up.

```{r pseudobulk_setup}
library(DESeq2)

pseudobulk_samples <- c("GSM9083544_FD16", "GSM9083546_FD28")

op_pb_input <- subset(op_fd_called, subset = sample %in% pseudobulk_samples)

# raw (unnormalized) counts required for DESeq2
counts_raw <- GetAssayData(op_pb_input, assay = "RNA", layer = "counts")

meta <- op_pb_input@meta.data
agg_group <- paste(meta$sample, meta$gnas_genotype_filtered, sep = "__")

# sum raw counts per patient x genotype pseudobulk group
pb_mat <- sapply(sort(unique(agg_group)), function(g) {
  Matrix::rowSums(counts_raw[, agg_group == g, drop = FALSE])
})
colnames(pb_mat) <- sort(unique(agg_group))

pb_mat  # 4 pseudobulk columns expected: FD16__MUT, FD16__WT, FD28__MUT, FD28__WT
```

```{r pseudobulk_deseq2}
coldata <- data.frame(
  pseudo_sample = colnames(pb_mat)
) %>%
  separate(pseudo_sample, into = c("patient", "genotype"), sep = "__", remove = FALSE) %>%
  column_to_rownames("pseudo_sample")

coldata$patient  <- factor(coldata$patient)
coldata$genotype <- factor(coldata$genotype, levels = c("WT", "MUT"))  # WT = reference

dds <- DESeqDataSetFromMatrix(
  countData = pb_mat,
  colData   = coldata,
  design    = ~ patient + genotype   # patient as blocking factor -> paired comparison
)

# filter out genes with essentially no signal across pseudobulk samples
dds <- dds[rowSums(counts(dds) >= 10) >= 2, ]

dds <- DESeq(dds)
pb_res <- results(dds, contrast = c("genotype", "MUT", "WT")) %>%
  as.data.frame() %>%
  rownames_to_column("gene") %>%
  arrange(padj)

head(pb_res, 20)
pb_res %>% filter(gene == "GNAS")

write.csv(pb_res, "Results/Osteoprogenitor_GNASmut_vs_WT_pseudobulk_DESeq2.csv", row.names = FALSE)
```

```{r pseudobulk_vs_percell}
# Cross-check: how many of the per-cell test's significant genes hold up in pseudobulk?
percell_sig  <- geno_results %>% filter(p_val_adj < 0.05) %>% pull(gene)
pseudobulk_sig <- pb_res %>% filter(!is.na(padj), padj < 0.05) %>% pull(gene)

length(percell_sig)
length(pseudobulk_sig)
length(intersect(percell_sig, pseudobulk_sig))

# Genes significant per-cell but NOT in pseudobulk are the ones most likely to be
# patient-confound artifacts rather than genuine genotype effects - treat with caution.
setdiff(percell_sig, pseudobulk_sig) %>% head(20)
```
